---
output: html_document
---
```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(VTatlas)
library(sf)
library(mapedit)
library(leaflet)
library(raster)
library(dismo)
```

```{r, echo = FALSE}
# Once this is automated it'll pull the sppVal from a console input when 
# running the job

##BSUB -q short
##BSUB -n 1
##BSUB -J "LepSpecies_MaxEnt[1-#NLEPS HERE]"
##BSUB -o "path/to/output/file on HPC_%I.out"
##BSUB -e "path/to/error/file on HPC_%I.err"
##BSUB -R "rusage[mem=18000]"
##BSUB -R "span[hosts=1]"
##BSUB -W 04:00
##BSUB -L /bin/bash
#module load R/3.5.1_gcc8.1.0; # load required modules
#module load gcc/8.1.0;
#module load gdal/3.0.4;
#module load geos/3.7.1;
#module load libpng/1.6.36;
#module load proj/6.0.0;
#Rscript path/to/R script $LSB_JOBINDEX

# this is one of the first lines in the R script to pull the $LSB_JOBINDEX
# into the R script to tell R which species to run the analysis for. 

# (sppVal <- as.numeric(commandArgs()[6]))

sppVal <- 1
```

```{r, warning = FALSE, message = FALSE}
### 1) Index for a species 

## Here I query the database for species within an order

a <- Sys.time()
LEPS <- queryAPI(query = list("order:Lepidoptera"))
Sys.time()-a

species <- unique(LEPS$scientificName)

n.species <- length(species)
```

There have been at least one observation of `r print(length(n.species))` species of
Lepidoptera (moths & caterpillars) in Vermont. 

### 2) Get number of observations (save value)

# `r species[sppVal]`

```{r,warning = FALSE, message = FALSE}
num_obs <- queryAPI(query = list(paste0("taxon_name:",species[sppVal])),
                    type = "TotalObs")
```
### 3) Grab data using API 
```{r,warning = FALSE, message = FALSE}
a <- Sys.time()
VAL_data <- queryAPI(query = list(paste0("taxon_name:",species[sppVal])),
                    type = "occurrences")
Sys.time()-a
```

### 4) Set up maxent analysis 
       a) just state of VT parameters?
       b) larger area and model in VT?

very basic use of MaxEnt model 
```{r, warning = FALSE, message = FALSE}
# make into spatial points data frame
VAL_data <- sf::st_as_sf(VAL_data, 
                         coords = c("decimalLongitude", "decimalLatitude"), 
                         crs = 4326, agr = "constant")

# remove duplicate locations 
# need to do that with sf 

# witholding a 25% sample for testing
fold <- kfold(VAL_data, k = 5)
testData <- VAL_data[fold == 1, ]
trainingData <- VAL_data[fold != 1, ]
```

```{r}
# THIS WILL ULTIMATELY PULL FROM A POSTGIS DATABASE 
towns <- sf::st_read("../../VermontAtlasofLife/VAL_GoogleDrive/Spatial_Layers/VAL Spatial Layers/VTtownsWGS84.shp")
```

### Gather environmental data

```{r}
# THESE ENV DATA NEED TO BE PROCESSED AND PLACED IN A DATABSE WITH THE SAME
# RESOLUTION, EXTENT, ETC. 
EnvData <- getData("worldclim", var="bio", res=0.5, lat = 43.79, lon = -72.63)
Elev <- getData("SRTM", lat = 43.79, lon = -72.63)
```

### Habitat suitability - Maxent

```{r, warning = FALSE, message = FALSE, results='hide'}
#fit the maxent model

# Keep only 
#BIO01 = Mean Annual Temp
#BIO06 = Min Temp Coldest Month
#BIO12 = Annual Precipitation
#BIO13 = Precipitation of Wettest Month
#BIO14 = Precipitation of Driest Month

EnvData <- dropLayer(EnvData,paste0("bio",c(2:5,7:11,15:19),"_13"))

EnvData <- suppressWarnings(mask(EnvData, towns))
Elev <- suppressWarnings(mask(Elev, towns))

EnvData <- suppressWarnings(crop(EnvData,towns))
```

```{r}
# Note - need to make elevation raster compatible with other layers @ 

# Run maxent with full dataset
MaxEnt_model <- maxent(x = EnvData, p = st_coordinates(trainingData))

# make predictions
suitHab <- predict(MaxEnt_model, EnvData)

# background data
bg <- randomPoints(EnvData, 1000) #background "pseudoabsences"

#simplest way to use 'evaluate'
AUCeval <- evaluate(MaxEnt_model, p=testData, a=bg, x=EnvData)
```

```{r, echo = FALSE}

set.breaks <- seq(0,1,0.01)
occcol <- colorRampPalette(c("white","gray88","blue","yellow","red"))(length(set.breaks))

#tiff("HabitatSuitability.tiff",res = 600, height = 3600, width = 3600, units = "px",compress = "lzw")
par(mar = c(0,0,0,0))
plot(st_geometry(towns),col = "gray88")
plot(suitHab,breaks = set.breaks,col = occcol,legend = FALSE,
     axes = FALSE, add = TRUE)
plot(st_geometry(towns), add = TRUE)
plot(suitHab,
     legend.only = TRUE,
    horizontal = TRUE,
    col= occcol,
    add = TRUE,
     breaks = set.breaks,
    smallplot = c(0.025,0.325,0.17,0.19),
     legend.width = 0.25,
     legend.shrink = 0.5,
     axis.args = list(at = seq(0,1,0.1),
                      las = 1,
                      labels = seq(0,1,0.1),
                      cex.axis = 1,
                      mgp = c(5,0.3,0)),
     legend.args = list(text = "Predicted\n Habitat Suitability", 
                      side = 3, 
                      font = 2, 
                      line = 0.5, 
                      cex = 1))
sppImg <- stack("G:/Spp_Images/Panopoda_rufimargo.jpg")
par(bty = "l", new = TRUE, fig = c(0.55,0.95,0.05,0.55))
plotRGB(sppImg, r=1,g=2,b=3,colNA = "black")
par(bty = "l",new = TRUE, fig = c(0.05,0.29,0.55,0.75),mar = c(1,1,1,1))
plot(AUCeval, 'ROC',type = "l",
     col = "black",lwd = 2,las = 1,main = "",cex.axis = 0.75,axes = FALSE,xaxt = "n")
axis(1,tck = -0.05,labels = seq(0,1,0.2),at = seq(0,1,0.2),cex.axis = 0.75,mgp = c(3, 0.2, 0.05),las = 1)
axis(2,las =2,tck = -0.05,labels = seq(0,1,0.2),at = seq(0,1,0.2),cex.axis = 0.75,mgp = c(3, 0.2, 0))
mtext(side = 1,text = "False positive rate",line = 1,cex = 0.75)
mtext(side = 2,text = "True positive rate",line = 1.2, cex = 0.75)
#dev.off()
#system("open HabitatSuitability.tiff")
```
       
### 6) Create a phenology figure 
```{r, warning = FALSE, message = FALSE}
VAL_data$observed <- 1

pheno <- tapply(VAL_data$observed,
                list(VAL_data$month),
                FUN = sum)

monthly_data <- array(0,c(1,12))
names(monthly_data) <- formatC(1:12, width = 2, flag = 0)

monthly_data[1,match(names(pheno),names(monthly_data))] <- pheno

par(bty = "l")
plot(monthly_data[1,]~c(1:12),
     las = 1, 
     ylab = "Number of observations",
     xlab = "Month",
     main = paste(species[sppVal]),
     pch = 19, col = "gray")


# NEED TO AUTOMATE THE SPAN VALUE SOMEHOW
# see http://r-statistics.co/Loess-Regression-With-R.html
smooth <- loess.smooth(x = 1:12,
                       y = monthly_data[1,],
                       span = 0.6)
points(smooth$y~smooth$x, type = "l")
```


### 7) Projections based on climate 
```{r, cache = TRUE, message = FALSE, warning = FALSE, results='hide'}
FutEnv <- getData("CMIP5", var="bio", 
                  rcp=85, year = 50, model = "AC",
                  res=2.5, lat = 43.79, lon = -72.63)

FutEnv <- dropLayer(FutEnv,paste0("ac85bi50",c(2:5,7:11,15:19)))

FutEnv <- suppressWarnings(mask(FutEnv, towns))
FutEnv <- suppressWarnings(crop(FutEnv,towns))

```

```{r, results='hide'}
# make predictions
names(FutEnv) <- names(EnvData)
FutHab <- predict(MaxEnt_model, FutEnv)

```

```{r, echo = FALSE}
FutHub_samp <- resample(FutHab,suitHab)

set.breaks <- seq(0,0.2,0.01)
occcol <- colorRampPalette(c("white","gray88","blue","yellow","red"))(length(set.breaks))

#tiff("HabitatSuitability.tiff",res = 600, height = 3600, width = 3600, units = "px",compress = "lzw")
par(mar = c(0,0,0,0))
plot(st_geometry(towns),col = "gray88")
plot(FutHub_samp,breaks = set.breaks,col = occcol,legend = FALSE,
     axes = FALSE, add = TRUE)
plot(st_geometry(towns), add = TRUE)
plot(FutHub_samp,
     legend.only = TRUE,
    horizontal = TRUE,
    col= occcol,
    add = TRUE,
     breaks = set.breaks,
    smallplot = c(0.025,0.325,0.17,0.19),
     legend.width = 0.25,
     legend.shrink = 0.5,
     axis.args = list(at = seq(0,1,0.1),
                      las = 1,
                      labels = seq(0,1,0.1),
                      cex.axis = 1,
                      mgp = c(5,0.3,0)),
     legend.args = list(text = "Future\n Habitat Suitability", 
                      side = 3, 
                      font = 2, 
                      line = 0.5, 
                      cex = 1))
sppImg <- stack("G:/Spp_Images/Panopoda_rufimargo.jpg")
par(bty = "l", new = TRUE, fig = c(0.55,0.95,0.05,0.55))
plotRGB(sppImg, r=1,g=2,b=3,colNA = "black")
par(bty = "l",new = TRUE, fig = c(0.05,0.29,0.55,0.75),mar = c(1,1,1,1))
plot(AUCeval, 'ROC',type = "l",
     col = "black",lwd = 2,las = 1,main = "",cex.axis = 0.75,axes = FALSE,xaxt = "n")
axis(1,tck = -0.05,labels = seq(0,1,0.2),at = seq(0,1,0.2),cex.axis = 0.75,mgp = c(3, 0.2, 0.05),las = 1)
axis(2,las =2,tck = -0.05,labels = seq(0,1,0.2),at = seq(0,1,0.2),cex.axis = 0.75,mgp = c(3, 0.2, 0))
mtext(side = 1,text = "False positive rate",line = 1,cex = 0.75)
mtext(side = 2,text = "True positive rate",line = 1.2, cex = 0.75)
#dev.off()
```





### Decline 

The difference between the two scenarios - currently CMIP5 - model: AC 8.5
```{r, echo = FALSE}

FutHub_samp <- resample(FutHab,suitHab)

diffHab <- FutHub_samp-suitHab



set.breaks <- seq(-1,1,0.01)
diffcol <- colorRampPalette(c("red","yellow","gray","white","gray","blue","forestgreen"))(length(set.breaks))

#tiff("HabitatSuitability.tiff",res = 600, height = 3600, width = 3600, units = "px",compress = "lzw")
par(mar = c(0,0,0,0))
plot(st_geometry(towns),col = "gray88")
plot(diffHab ,breaks = set.breaks,col = diffcol,legend = FALSE,
     axes = FALSE, add = TRUE)
plot(st_geometry(towns), add = TRUE)
plot(diffHab,
     legend.only = TRUE,
    horizontal = TRUE,
    col= diffcol,
    add = TRUE,
     breaks = set.breaks,
    smallplot = c(0.025,0.325,0.17,0.19),
     legend.width = 0.25,
     legend.shrink = 0.5,
     axis.args = list(at = seq(-1,1,0.25),
                      las = 1,
                      labels = seq(-1,1,0.25),
                      cex.axis = 1,
                      mgp = c(5,0.3,0)),
     legend.args = list(text = "Future\n Habitat Suitability", 
                      side = 3, 
                      font = 2, 
                      line = 0.5, 
                      cex = 1))
sppImg <- stack("G:/Spp_Images/Panopoda_rufimargo.jpg")
par(bty = "l", new = TRUE, fig = c(0.55,0.95,0.05,0.55))
plotRGB(sppImg, r=1,g=2,b=3,colNA = "black")
par(bty = "l",new = TRUE, fig = c(0.05,0.29,0.55,0.75),mar = c(1,1,1,1))

```
